---
title: "Multivariate Linear Regression.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
For the porpouse of trying to explain our hypotesis, we are going to carry on a multivariate linear regression. 
```{r cars}
mydata = read.csv("Cancer_final.csv")
data3 = na.omit(mydata)

```
Now let's regress cancer against all the covariates that we have.
```{r}
fit = lm( Cancer ~ Alcool + GDP + Obesity + Age + Smoking, data = data3)
fit
summary(fit)


```
From just these  results, we can see various things: the value of the Adjusted R-Squared is satisfying, and then this model fits pretty well the data. Also, all the variables seems to explain pretty well the number of cancer cases in each country.
On the other hand, if we try for example just to consider data coming from just one year, the results are pretty surprising

```{r}
data4 = subset(data3, Year == "2005")
fit2 = lm( Cancer ~ Alcool + GDP + Obesity + Age + Smoking, data = data4)
fit2
summary(fit2)

#fin = subset(mydata, select = c(Entity,Year,Cancer,Smoking))
#fan = na.omit(fin)
#fit2 = lm (Cancer ~ Smoking, data= fin)
#fit2
#summary(fit2)
#plot(Cancer ~ Smoking, data=fin)
#abline(fit2, col="blue", lwd=2)
```
In this case it seems that some of the variables do not correlate well! We can try the conclusion that the subset of the dataset we are considering ( just one year ) is not satisfying to take some satisfactory conclusion.
```{r}
fin = subset(mydata,Year=="2005", select = c(Entity,Cancer,Smoking))
fan = na.omit(fin)
fit3 = lm (Cancer ~ Smoking, data= fin)
fit3
summary(fit3)
plot(Cancer ~ Smoking, data=fin)
abline(fit3, col="blue", lwd=2)
```
Plotting just a simple linear regression, we can see and find also from the p-value that there is a strong correllation between cancer and smoking. Anyway, for what we can see from the R-Squared value, the model does not explain pretty well our data.

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
